---
title: "Final Project"
authors: "Kim Mincin & Drew Insley"
due date: "April 30, 2019"
output: html_notebook
---

## Front Matter

```{r message=FALSE, warning=FALSE}
# clear enviornment
rm(list = ls())

# download packages
library(mdsr)
library(rvest)
library(readr)
library(tidyr)
library(rgeos)
library(rworldmap)

# read data
page <- "http://worldpopulationreview.com/countries/suicide-rate-by-country/"
xpath <- '//*[@id="__next"]/div/div[1]/div[2]/div[1]/div/div[2]/div/div/table'
table_list <- 
  page %>%
  read_html() %>%
  html_nodes(xpath = xpath) %>%
  html_table(fill = TRUE)
Suicide <- table_list[[1]]

Happiness <- read.csv("2017_Happiness.csv")

Happy_Suicide <- 
  Happiness %>%
  left_join(Suicide, by = c("Country" = "X1"))

# user - defined functions
fitted_model <- function(a, data) {
  # fits a linear model based on provided estimates

  a[1] + data$GDP_per_capita * a[2]
}

resid_measurement <- function(mod, data) {
  # computes the root mean squared of a linear model based on provided estimates
  
  diff <- data$happiness_score - fitted_model(mod, data)
  sqrt(mean(diff ^ 2))
}
```

## Intro
To utilize our data science skills for social good we choose to investigate happiness levels across the world. Our first data set, The World Happiness Report (2017_Happiness.csv), scores each countries happiness using data from the Gallup World Poll. The dataset was released at a UN event celebrating International Day of Happiness. The other variables included in the data set can be used as predictors for happiness. These predictors once understood could be used to influence public policy and assess countries progress. Our second data set provided the latitude and longitude of each country so that happiness could be evaluated on a regional basis as well as locally. The goal of our project was to determine what factors have the greatest influence on a countries happiness so that leaders can better target how to improve the well being of it’s citizens.

## Project

```{r}
# clean data
Happy_Suicide_Clean <-
  Happy_Suicide %>%
  rename(country = Country,
         happiness_rank = Happiness.Rank,
         happiness_score = Happiness.Score,
         maximum_happiness = Whisker.high,
         minimum_happiness = Whisker.low,
         GDP_per_capita = Economy..GDP.per.Capita.,
         family = Family,
         life_expectancy = Health..Life.Expectancy.,
         freedom = Freedom,
         generosity = Generosity,
         corruption = Trust..Government.Corruption.,
         suicides = X2,
         suicide_rank = X3) %>%
  mutate(suicides = as.numeric(gsub("suicides per 100k", " ", suicides)),
         suicide_rank = as.numeric(suicide_rank))

Happy_Suicide_Clean[12] <- NULL

Happy_Suicide_Clean
```

```{r}
# create density plots
Happy_Suicide_Clean_Top_25 <- 
  Happy_Suicide_Clean %>%
  arrange(happiness_rank) %>%
  head(25)

Happy_Suicide_Clean_Bottom_25 <-
  Happy_Suicide_Clean %>%
  arrange(desc(happiness_rank)) %>%
  head(25)

Happy_Suicide_Clean_Top_25 %>%
  ggplot(aes(x = GDP_per_capita)) +
  geom_density(fill = "dark red", alpha = .3) +
  geom_density(aes(x = GDP_per_capita), data = Happy_Suicide_Clean_Bottom_25, fill = "dark blue", alpha = .3) +
  ggtitle("Density Plots of Top/Bottom 25 Countries GDP Per Capita") +
  theme(plot.background = element_rect(fill = 'grey'),
        plot.title = element_text(size = 17.5, lineheight = .8, vjust = 1, hjust = .5, family = "", face = "bold"),
        panel.background = element_rect(fill = 'grey'),
        panel.grid.major = element_line(color = "white"),
        panel.grid.minor = element_line(color = "grey"))
```

We began by showing a density plot for GDP_per_capita for the top and bottom 25 happiest countries.

```{r}
# find best coefficient estimates
Happy_Best <- optim(par = c(0, 0), fn = resid_measurement, data = Happy_Suicide_Clean)
Happy_Best$par
```
We started by finding the optim parameters for intercept and slope for a model predicting happiness score for each country based on GDP per capita.

```{r}
# create optimal linear model
Happy_Suicide_Clean %>%
  ggplot(aes(x = GDP_per_capita, y = happiness_score)) +
  geom_abline(intercept = 3.203957, slope = 2.183670, color = "dark red", 
                 linetype = "solid", size = .5) +
  geom_point(color = "dark blue") +
  xlab("GDP Per Capita") +
  ylab("Happiness Score") +
  ggtitle("Predicting Happiness Score from GDP Per Capita") +
  annotate("text", x = .4, y = 7, label = "y = 2.183670x + 3.203957", size = 7, color = "dark green") +
  theme(plot.background = element_rect(fill = 'grey'),
        plot.title = element_text(size = 17.5, lineheight = .8, vjust = 1, hjust = .5, family = "", face = "bold"),
        panel.background = element_rect(fill = 'grey'),
        panel.grid.major = element_line(color = "white"),
        panel.grid.minor = element_line(color = "grey"))
```
This graph showcases the results of our optim function for happiness score.

```{r}
# create optimal residual plot
mod_values <- 3.203957 + (2.183670 * Happy_Suicide_Clean$GDP_per_capita)

Happy_Dist <-
  Happy_Suicide_Clean %>%
  select(country, happiness_score) %>%
  mutate(model_values = mod_values,
         residuals = happiness_score - model_values)

Happy_Dist %>%
  ggplot(aes(x = model_values, y = residuals)) +
  geom_point(color = 'dark blue') +
  geom_hline(yintercept = 0, color = 'dark red') +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot") +
  annotate("text", x = 6.65, y = 1.75, label = "RMS Error = 0.66", size = 7.5, color = "dark green") +
  theme(plot.background = element_rect(fill = 'grey'),
        plot.title = element_text(size = 20, lineheight = .8, vjust = 1, hjust = .5, family = "", face = "bold"),
        panel.background = element_rect(fill = 'grey'),
        panel.grid.major = element_line(color = "white"),
        panel.grid.minor = element_line(color = "grey"))
```
There is no distinct pattern among the model so the residual plot shows that the linear model is a good fit for our data.

```{r}
# calculate RMS error
sqrt(mean(Happy_Dist$residuals ^ 2))
```
The root mean squared error for the optimal line is .657.

```{r}
Happy_Min <-
  Happy_Suicide_Clean_Top_25 %>%
  select(country, minimum_happiness) %>%
  spread(key = country, value = minimum_happiness)

Happy_Max <- 
  Happy_Suicide_Clean_Top_25 %>%
  select(country, maximum_happiness) %>%
  spread(key = country, value = maximum_happiness)

Happy_Mix_Max <-
  Happy_Min %>%
  rbind(Happy_Max)

apply(X = Happy_Mix_Max, FUN = mean, MARGIN = 2)
```

We looked at a measurement for the average happiness for the top 25 happiest countries, and began to notice that countries in the same region has similar levels of happiness. To further explore this, we decided to create a dendogram to examine the similarities between countries.

```{r}
# dendogram of top 25 happiest countries
Happy_Suicide_Clean_2 <- Happy_Suicide_Clean[-c(1)]

Happy_Suicide_Clean_2 <-
  Happy_Suicide_Clean_2 %>%
  select(happiness_rank, GDP_per_capita, life_expectancy, generosity, freedom) %>%
  arrange(happiness_rank) %>%
  head(25)
  
Happy_std <-
  scale(Happy_Suicide_Clean_2) %>%
  as.data.frame()

Happy_dist <- 
  dist(Happy_std)

Happy_dendo <-
  Happy_dist %>%
  hclust(method = "complete")

labels <- Happy_Suicide_Clean_Top_25$country

Happy_dendo %>%
  plot(cex = 0.9, labels = labels, lwd = 2,
       main = "Top 25 Happiest Countries Dendogram")
```
The dendogram shows the dissimularity among the top 25 happiest countries. Interestingly the countries are grouped together by region. 

```{r}
# show happiness levels on a map
map <- getMap(resolution = "low")
coordinates <- gCentroid(map, byid = TRUE)
Lat_Long <- 
  coordinates %>%
  as.data.frame()

Lat_Long_Clean <-
  setNames(cbind(rownames(Lat_Long), Lat_Long, row.names = NULL), 
         c("country", "latitude", "longitude"))

Happy_Places <-
  Happy_Suicide_Clean %>%
  left_join(Lat_Long_Clean) %>%
  select(country, latitude, longitude, happiness_score) %>%
  filter(latitude != "NA") %>%
  mutate(happiness_score = ifelse(happiness_score >= mean(happiness_score), "Top 50%", "Bottom 50%"))

Happy_Places %>%
  ggplot(aes(x = latitude, y = longitude)) +
  geom_point(aes(color = happiness_score), size = 3) +
  geom_point(aes(x = -75, y = 40), color = "cyan3", size = 3) + 
  annotate("text", x = -75, y = 45, label = "United States", size = 4, color = "white")  +
  geom_hline(yintercept = 0, color = 'dark red', linetype = 'dotted') +
  geom_vline(xintercept = 0, color = 'dark red', linetype = "dotted") +
  xlab("Latitude") +
  ylab("Longitude") +
  ggtitle("Happiness on the Map") +
  theme(plot.background = element_rect(fill = 'deepskyblue3', color = "deepskyblue3"),
        plot.title = element_text(size = 20, lineheight = .8, vjust = 1, family = "", face = "bold"),
        panel.background = element_rect(fill = 'deepskyblue3'),
        panel.grid.major = element_line(color = "deepskyblue3"),
        panel.grid.minor = element_line(color = "deepskyblue3"),
        legend.background = element_rect(fill = "deepskyblue3"),
        legend.key = element_rect(fill = "deepskyblue3", color = NA))
```

To show that the split of happiness was related to region we decided to graph the countries location along with happiness score. 
